---
title: 'Patricia: Grupos Funcionales'
author: "Jorge Cornejo"
date: "10/25/2018"
output:
  html_document:
    highlight: haddock
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 4
    toc_float: yes
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(results = 'asis')
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width  = 9)
knitr::opts_chunk$set(fig.height = 6)
```

**Install rFishBase package**: This only need to be done once.
```{r eval=F}
remotes::install_github("ropensci/rfishbase")
```

```{r echo=F, warning=FALSE, results='hide', message=F}
rm(list=ls())
library(readxl)
library(dplyr)
library(tidyr)
library(DT)
require(rfishbase)
require(FD) #This calculates functional diversity
```


# Read data and obtain Species Names

## Data in Capturas

```{r}
capturas <- read_excel('Data/Solicitud Patricia Faundez AH002T0002612.xlsx', sheet = 'Capturas')

spNames <- capturas %>% 
  select(comun, cientifico) %>% 
  dplyr::group_by(comun, cientifico) %>% 
  dplyr::summarise(n=n())

datatable(head(spNames))
```

## Especies Consideradas

```{r}
spp <- read_excel('Data/Solicitud Patricia Faundez AH002T0002612.xlsx', sheet = 'Especies consideradas')

spp <- spp %>% 
  select(COD_ESP, nombre_común, simbolo)
  
names(spp) <- c("COD_ESP", "nombre_comun", "simbolo" )

datatable(head(spp))
```

## Merge the tables

```{r}
spp <- left_join(spp, spNames, by=c("nombre_comun"="comun"))
datatable(spp)
```

## Get information from FishBase

Fist we remove the non-fishes species. 

~~Then we replace the species name **_Granadero aconcagua_** for **_Coelorinchus aconcagua_**, which is the names that it is in FishBase.~~

```{r}
spp <- spp %>%
            filter(!COD_ESP %in% c(48, 52, 561, 35, 136, 135, 132, 874))
```


```{r cache=TRUE}
fishbaseInfo <- species(spp$cientifico)
fishbaseInfo <- left_join(spp, fishbaseInfo, by=c("cientifico"="Species"))
temp <- ecology(spp$cientifico, fields=c("SpecCode", "DietTroph")) %>% 
    rename(TL=DietTroph)
fishbaseInfo <- left_join(fishbaseInfo, temp, by=c("SpecCode"="SpecCode"))

datatable(fishbaseInfo[1:5,1:15])
```

## Cleanup the columns

```{r}
fishbaseInfo <- fishbaseInfo %>%
              select(-n, -SpeciesRefNo, -FBname, -PicPreferredName, -PicPreferredNameM, -PicPreferredNameF,
               -PicPreferredNameJ, -Source, -AuthorRef, -Remark, -TaxIssue, -Fresh, -Brack, -Saltwater, -MigratRef,
               -Pic, -PictureFemale, -LarvaPic, -EggPic, -PriceCateg, -PriceReliability, -Remarks7, -UsedasBait, -BaitRef,
               -Aquarium, -AquariumFishII, -AquariumRef, -GameFish, -GameRef, -Dangerous, -DangerousRef, -Electrogenic,
               -ElectroRef, -Complete, -GoogleImage, -Emblematic, -Entered, -DateEntered, -Modified, -DateModified,
               -Expert, -DateChecked, -TS)
```

## Write a CSV file
```{r}
write.csv(fishbaseInfo, file="Data/Derived/Species.csv")
```

# Funcionatal diversity

## Clean data from before

And then read the data and keep only the traits of interest.
```{r}
#rm(list=ls())
data <- read.csv(file="Data/Derived/Species.csv")
cData <- data %>% 
  select(cientifico, TL, DemersPelag, DepthRangeShallow, DepthRangeDeep, Length, BodyShapeI) %>% 
  arrange(cientifico) %>% 
  filter(cientifico != "Coelorhynchus aconcagua")
row.names(cData) <- cData$cientifico

capturas <- read_excel('Data/Solicitud Patricia Faundez AH002T0002612.xlsx', sheet = 'Capturas')
```

## Sum catches by Species and Year

```{r}
yearCatch <- capturas %>% 
  separate(proy, c("year", "nn"), sep="-", remove = F) %>% 
  separate(year, c("year", "nn"), sep="_", remove = F) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(year = ifelse(year>1000, year, year+1900)) %>% 
  group_by(year, cientifico) %>% 
  summarise(totalCatch_kg = sum(captkg, na.rm=T)) %>% 
  filter(cientifico %in% cData$cientifico) %>% 
  arrange(cientifico, year)
```

## Turn into Wide format

### Abundance Matrix
This values are calculates as the sum of catches by species on a giver year.
Then year is turn int a **row.name** and is drop from the columns.

_Coelorhynchus aconcagua_ does not have any data, so we remove it from the table.

```{r results="html"}

abundances <- yearCatch %>% 
  filter(cientifico != "Coelorhynchus aconcagua") %>% 
  spread(cientifico, totalCatch_kg)

row.names(abundances) <- abundances$year
abundances <-  abundances[,-1] %>% 
  replace(., is.na(.), 0)

cData <- select(cData, -cientifico)
write.csv(cData, file="Data/Derived/TraitsSP.csv")
print(row.names(cData))
```

## FD of the species 


```{r}
test <- dbFD(cData, cor = "cailliez", a=abundances) #, cor="cailliez")
datatable(data.frame(test[1:8]))
datatable(test$CWM)
```


# Now working with Chilean Landings

## Load and keep only Bottom Trawling
```{r cache=TRUE}
landings <- read.csv(file="Data/Industrial_Chile.csv", stringsAsFactors = F)
```

```{r}
trawling <- landings %>% 
  filter(gear == "ARRASTRE") %>% 
  group_by(year, stock) %>% 
  summarise(totalCatch = sum(landings, na.rm=T))
```

Now prepare the **spNames** _dataset_ to get spepecies names for the species on **trawling** _dataset_.

```{r}
spNames <- spNames %>% 
  mutate(common=toupper(comun))
```

Merge **speNames** with **trawling** to add Species names.

```{r}

spTrawling <- landings %>% 
  filter(gear == "ARRASTRE") %>% 
  group_by(stock) %>% 
  summarise(n = n()) %>% 
  separate(stock, c("comun1", "comun2"), sep = " O ") %>% 
  separate(comun1, c("comun1", "comun3"), sep = " / ") %>% 
  mutate(comun2=ifelse(!is.na(comun2), comun2, comun3)) %>% 
  select(-comun3, n)


spTrawling <- left_join(spTrawling, spNames, by=c("comun1"="common"))
spTrawling <- left_join(spTrawling, spNames, by=c("comun2"="common"))

spTrawling <- spTrawling %>% 
  mutate(spName = ifelse(!is.na(cientifico.x), cientifico.x, cientifico.y)) %>% 
  rename(common = comun1) %>% 
  select(common, spName)

spTrawling$spName <- ifelse(spTrawling$common ==  "MERLUZA COMUN", "Merluccius gayi gayi", spTrawling$spName)
```

There are many species **without** scientific names. I will move on for now, removing them!!

```{r}
spTrawlingWithScName <- spTrawling %>% 
  filter(!is.na(spName))
```

## Get data from FishBase

### Traits Table

Get FishBase data and then drop those that are not fishes of FB does not have data on them.
Finally we only keep the columns that are needed for the analysis.

This table are the input **traits** to the _FD_ functions.
```{r cache=TRUE}
fbSpTrawling <- species(spTrawlingWithScName$spName)

fbSpTrawling <- cbind(spTrawlingWithScName, fbSpTrawling)

fbSpTrawling$TL <- ecology(fbSpTrawling$spName, fields=c("DietTroph"))$DietTroph  ## This get the Trophic Level from FB


fbSpTrawling <- fbSpTrawling %>% 
  select(spName, TL, DemersPelag, DepthRangeShallow, DepthRangeDeep, Length, BodyShapeI) %>% 
  filter(!spName %in% c("Loligo gahi", "Rapana (Chorus) giganteus", "Lithodes santolla", "Haliporoides diomedeae", 
                         "Homolaspis plana", "Mursia gaudichaudii", "Dosidicus gigas", "Cervimunida johni",
                         "Pleuroncodes monodon", "Enteroctupus megalociathus")) %>% 
  arrange(spName)
  
row.names(fbSpTrawling) <- fbSpTrawling$spName
traits <- fbSpTrawling %>% 
  select(-spName)

datatable(traits)
```

### Now format the landings

Remove all the stock whitout a Scientific name!

```{r}
trawling2 <- left_join(trawling, spTrawlingWithScName, by=c("stock"="common"))
abundances <- trawling2 %>% 
  filter(!is.na(spName), spName %in% row.names(traits)) %>% 
  select(-stock) %>% 
  spread(spName, totalCatch)

row.names(abundances) <- abundances$year
abundances <-  abundances[,-1] %>% 
  replace(., is.na(.), 0)
```

# FD derived from landings

For some weir reason there are more species on traits than on abundances. I'm removing
all species that are not in abundances table.

```{r}
traits$sp <- row.names(traits)
traits <- traits %>% 
  filter(sp %in% names(abundances)) 

traits <- traits %>% 
  mutate(DemersPelag = as.factor(DemersPelag), BodyShapeI = as.factor(BodyShapeI ))

row.names(traits) <- traits$sp

traits <- select(traits, -sp)
write.csv(traits, file="Data/Derived/Traits_Catches.csv")
```


```{r}
test <- dbFD(traits, cor = "cailliez", a=abundances) 
datatable(data.frame(test[1:8]))
datatable(test$CWM)
```

