---
title: 'Patricia: Grupos Funcionales'
author: "Jorge Cornejo"
date: "10/25/2018"
output:
  html_document:
    highlight: haddock
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 4
    toc_float: yes
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(results = 'asis')
knitr::opts_chunk$set(warning = TRUE)
knitr::opts_chunk$set(message = TRUE)
knitr::opts_chunk$set(fig.width  = 9)
knitr::opts_chunk$set(fig.height = 6)
```

**Install rFishBase package**: This only need to be done once.
```{r eval=F}
remotes::install_github("ropensci/rfishbase")
```

```{r echo=F, warning=FALSE, results='hide', message=F}
rm(list=ls())
library(readxl)
library(dplyr)
library(tidyr)
library(DT)
require(rfishbase)
require(FD) #This calculates functional diversity
```

# Read data and obtain Species Names

## Data in Capturas

```{r}
capturas <- read.csv('Data/WCsurvey2.csv', stringsAsFactors = F)

# Three bio geographic regions were defined using published literature suggesting potentially significant faunal changes within these regions (Gabriel 1982, Williams & Ralston 2002, Francis et al. 2009) The biogeographic regions are: south of Cape Mendocino, Cape Mendocino to Cape Blanco, north of Cape Blanco. For example abundant species as Pacific Cod '

sppWC <- capturas %>%
  filter(latitude_dd >34) %>% #consider 34 lat S that includes Santa Barbara port
  filter(latitude_dd <40) %>%
  filter(!year %in% c("1977","1980","1983","1986", "1989","1992", "1995", "1998", "1999", "2000", "2001", "2002")) %>%
  filter(year > 2013) %>%  #I am analysing 10 years in WC 
  select(scientific_name) %>%
  group_by(scientific_name) %>%
  summarise(n=n()) %>%
  rename (cientifico = scientific_name)

sppWC2 <- capturas %>%
  filter(latitude_dd >34) %>% #consider 34 lat S that includes Santa Barbara port
  filter(latitude_dd <40) %>%
  filter(!year %in% c("1977","1980","1983","1986", "1989","1992", "1995", "1998", "1999", "2000", "2001", "2002")) %>%
  filter(year > 2013)  #I am analysing 10 years in WC 

  survey_min_depth <- sppWC2 %>% #looking at what depth will be more appropiate to cut the landings data based on what is sampled in surveys (the percentage of species in surveys at 20, 3, 40, 50, 60, 70, 80, 90, 100)
  filter(depth_m < 70) %>% # trawling' hauls started al min depth of 60.7. Thus, in landings I deleted all species with depth ranges shallowed than 60.
  select(depth_m) %>% 
    mutate(min(depth_m))

```


## Get information from FishBase

Fist we remove the non-fishes species.

```{r cache=TRUE}
fishbaseInfo1 <- species(sppWC$cientifico)
fishbaseInfo2 <- left_join(sppWC, fishbaseInfo1, by=c("cientifico"="Species")) %>%
  filter(SpecCode != "NA")

sppDeepWC2 <- inner_join(sppWC, fishbaseInfo2, by = "cientifico") # getting traits from intersection with surveys traits

temp <- ecology(sppDeepWC2$cientifico, fields=c("SpecCode", "DietTroph")) %>%
    rename(TL=DietTroph)
fishbaseInfo <- left_join(fishbaseInfo2, temp, by=c("SpecCode"="SpecCode")) #171 sp

```

## Cleanup the columns
```{r}
cData2 <- fishbaseInfo %>%
  select(cientifico, TL, DemersPelag, DepthRangeShallow, DepthRangeDeep, Length, BodyShapeI) %>%
  arrange(cientifico)

#what_these_are_different <- anti_join(cData2, cData)#Caristius macropus, is because this sp doesn't have abundances in this area and cData have species from the survey that have abundances

# cData <- fishbaseInfo %>%
#   select(cientifico, TL, DemersPelag, DepthRangeShallow, DepthRangeDeep, Length, BodyShapeI) %>%
#   arrange(cientifico)

#cData1 <- cData[-58,]
# #cData1 <- read.csv('Data/Derived/WC_survey_traits_0_600_35_40_final.csv') %>%
#   mutate(cientifico = as.character(cientifico)) %>%
#   mutate(TL = as.character(TL)) %>%
#   mutate(DemersPelag = as.character(DemersPelag)) %>%
#   mutate(DepthRangeShallow = as.character(DepthRangeShallow)) %>%
#   mutate(DepthRangeDeep = as.character(DepthRangeDeep)) %>%
#   mutate(Length = as.character(Length)) %>%
#   mutate(BodyShapeI = as.character(BodyShapeI))

#cData2 <- cData1
 
#row.names(cData) <- cData$cientifico

#cData2 <- read.csv('Data/Derived/WCsurveytraits2.csv', stringsAsFactors = F)
#filter(cientifico != "Gadus macrocephalus" & TL != 4.16) #3.78)
#cData <- cData[cData$cientifico!="Gadus macrocephalus" & cData$TL != 3.78,]
```

## Sum catches by Species and Year


```{r}
cD <- fishbaseInfo %>%
  select(cientifico, TL, DemersPelag, DepthRangeShallow, DepthRangeDeep, Length, BodyShapeI) %>%
  arrange(cientifico)

cData3 <- cD %>%  
  select(cientifico)

yearCatch1 <- capturas %>%
 # filter(depth_m < 600) %>% #AS IF THIS SHOULD BE A PROBLEM WHEN COMPARING WITH CHILE
  filter(latitude_dd >34) %>%
  filter(latitude_dd <40) %>%
  filter(!year %in% c("1977","1980","1983","1986", "1989","1992", "1995", "1998", "1999", "2000", "2001", "2002")) %>%
  filter(year < 2013) %>% #selecting years for the current trawling survey sampling 2003 - 2013(cite)
  select(cpue_kg_per_ha_der, scientific_name, year) %>%
  rename(captkg = cpue_kg_per_ha_der) %>%
  rename(cientifico = scientific_name)

yearCatch <- yearCatch1 %>%
  select(year, cientifico, captkg) %>%
  group_by(year, cientifico) %>%
  summarise(totalCatch_kg = sum(captkg, na.rm=T)) %>%
  #filter(cientifico %in% cData3$cientifico) %>%
  arrange(cientifico, year)  



yearCatch <- inner_join(yearCatch, cData3, by = "cientifico") %>%
  select(year, cientifico, totalCatch_kg)

# Deleting species with catch = 0 kg. in WCsurvey 

cD <- fishbaseInfo %>%
  select(cientifico, TL, DemersPelag, DepthRangeShallow, DepthRangeDeep, Length, BodyShapeI) %>%
  arrange(cientifico)

abundance2 <- yearCatch %>% 
  select(cientifico) %>%
  group_by(cientifico) %>%
  summarise(n=n())
  #distinct(cientifico, .keep_all = TRUE)

cData <- left_join(abundance2, cD, by = 'cientifico') %>% 
  select(-n)

```



```{r}
cD3 <- fishbaseInfo %>%
  select(cientifico, TL, DemersPelag, DepthRangeShallow, DepthRangeDeep, Length, BodyShapeI) %>%
  arrange(cientifico)

abundance3 <- yearCatch %>% 
  select(cientifico) %>%
  group_by(cientifico) %>%
  summarise(n=n())
  #distinct(cientifico, .keep_all = TRUE)

cD4 <- left_join(abundance3, cD3, by = 'cientifico') %>% 
  select(-n)

# Survey_completed <- read.csv('Data/Derived/WC_survey_traits_0_600_35_40_final.csv', stringsAsFactors = F) %>% # this is the survey data completed with searched information, which for most sp were found in fishbase.
#   mutate(TL = as.character(TL)) %>%
#   mutate(DepthRangeShallow = as.character(DepthRangeShallow)) %>%
#   mutate(DepthRangeDeep = as.character(DepthRangeDeep)) %>%
#   mutate(Length = as.character(Length))

Survey_completed <- read_xls('Data/Derived/test1.xls') %>% 
  select(-...1)


Survey <- cD4 %>%   
  select(cientifico)

cData <- left_join(Survey, Survey_completed, by ='cientifico')

cData <- cData %>%
  mutate(TL = as.double(TL)) %>%
  mutate(DepthRangeDeep = as.numeric(DepthRangeDeep)) %>%
  mutate(DepthRangeShallow = as.numeric(DepthRangeShallow))%>%
  mutate(Length = as.numeric(Length))

write.csv(cData, file="Data/Derived/WC_Survey_traits_completed.csv")# exported to fill NA's with info from scientific sources

```


## Turn into Wide format

### Abundance Matrix
This values are calculates as the sum of catches by species on a giver year.
Then year is turn int a **row.name** and is drop from the columns.

_Coelorhynchus aconcagua_ does not have any data, so we remove it from the table.

```{r results="html"}
abundances <- yearCatch %>%
  spread(cientifico, totalCatch_kg) %>% 
  select(-'Caristius macropus')

row.names(abundances) <- abundances$year
abundances <-  abundances[,-1] %>%
  replace(., is.na(.), 0)


temp <- as.data.frame(colSums(abundances))
temp$names <- row.names(temp)
temp2 <- temp[temp$`colSums(abundances)`==0,]

abundances <- abundances %>%
  select(-c(temp2$names))

cData <- cData %>% 
  filter(cientifico!='Caristius macropus') %>% 
  filter(!cientifico %in% temp2$names ) %>%
  arrange(cientifico) %>% 
  mutate(DemersPelag=as.factor(DemersPelag), BodyShapeI = as.factor(BodyShapeI))


nameList <- cData$cientifico
cData <- cData[,-1]
row.names(cData) <- nameList

testFDcommunity <- dbFD(cData, corr = "cailliez", a=abundances)
datatable(data.frame(testFDcommunity[1:8]))
datatable(testFDcommunity$CWM)

# write.csv(cData, file="Data/Derived/WC_survey_traits.csv")
# write.csv(abundances, file="Data/Derived/WC_survey_abundances.csv")
# write.csv(testFDcommunity, file="Data/Derived/WC_survey_FD.csv")
```


```{r results="html"}
abundances <- yearCatch %>%
  spread(cientifico, totalCatch_kg)

row.names(abundances) <- abundances$year
abundances <-  abundances[,-1] %>%
  replace(., is.na(.), 0)


temp <- as.data.frame(colSums(abundances))
temp$names <- row.names(temp)
temp2 <- temp[temp$`colSums(abundances)`==0,]

abundances <- abundances %>%
  select(-c(temp2$names))

Surveytraits_completed - Surveytraits_completed %>%
  mutate(TL = as.double(TL)) %>%
  mutate(DepthRangeDeep = as.numeric(DepthRangeDeep)) %>%
  mutate(Length = as.numeric(Length))

Surveytraits_completed <- Surveytraits_completed %>%
  filter(!cientifico %in% temp2$names ) %>%
  arrange(cientifico)

nameList <- Surveytraits_completed$cientifico
Surveytraits_completed <- Surveytraits_completed[,-1]
row.names(Surveytraits_completed) <- nameList

Surveytraits_completed2 - Surveytraits_completed %>%
  mutate(TL = as.double(TL)) %>%
  mutate(DepthRangeDeep = as.numeric(DepthRangeDeep)) %>%
  mutate(Length = as.numeric(Length))


testFDcommunity <- dbFD(Surveytraits_completed, corr = "cailliez", a=abundances)
datatable(data.frame(testFDcommunity[1:8]))
datatable(testFDcommunity$CWM)



write.csv(Surveytraits_completed, file="Data/Derived/WC_survey_traits.csv")
write.csv(abundances, file="Data/Derived/WC_survey_abundances.csv")
write.csv(testFDcommunity, file="Data/Derived/WC_survey_FD.csv")
```

# Now working with Landings, check illegal fishing?

